Networked Character Devices

### 2011-12-09 report/presentation.tex
presentation: center images
presentation: remove code sample
presentation: add protocol architecture frame
presentation: use fixed font for some terms, vimspell
presentation: Most of Protocol and Client subsections

\begin{frame}[c]{Very High Level}
  \includegraphics[width=0.8\textwidth]{arch-01.png}

@@

\begin{frame}[c]{High Level}
  \includegraphics[width=0.8\textwidth]{arch-02.png}

@@

\begin{frame}[c]{Client Medium Level}
  \includegraphics[width=0.4\textwidth]{arch-03.png}

  \begin{itemize}
    \item<1-> By default, devices are exposed under /dev/netchar
    \item<2-> In the future, we should be able to choose any name
  \end{itemize}
\end{frame}

\begin{frame}[c]{Server Medium Level}
  \includegraphics[width=0.8\textwidth]{arch-02.png}
  \begin{itemize}
    \item<1-> Listen on a specified port
    \item<2-> Receive requests from client
    \begin{itemize}
      \item<2-> open,close,read,write,...
    \end{itemize}
    \item<3-> Can actually export \textit{any} file
    \item<4-> Crashes ungracefully when the client disconnects
  \end{itemize}
\end{frame}

@@

%---Code Sample Slide---%
\begin{frame}{\bf Code Sample}

\lstinputlisting[
  caption={Listing Code Sample},
  label=lst:codeSample.c]{codesample.c}

@@

% vim: set sw=2 ts=2 et : %

>>>>

\begin{frame}[c]{Very High Level}
  \begin{center}
    \includegraphics[width=0.8\textwidth]{arch-01.png}
  \end{center}

@@

\begin{frame}[c]{High Level}
  \begin{center}
    \includegraphics[width=0.8\textwidth]{arch-02.png}
  \end{center}

@@

\begin{frame}[c]{Client Medium Level}
  \begin{center}
    \includegraphics[width=0.4\textwidth]{arch-03.png}
  \end{center}

  \begin{itemize}
    \item<1-> By default, devices are exposed under \texttt{/dev/netchar}
    \item<2-> In the future, we should be able to choose any name
  \end{itemize}
\end{frame}

\begin{frame}[c]{Server Medium Level}
  \begin{center}
    \includegraphics[width=0.8\textwidth]{arch-02.png}
  \end{center}
  \begin{itemize}
    \item<1-> Listen on a specified port
    \item<2-> Receive requests from client
    \begin{itemize}
      \item<2-> \texttt{open()},\texttt{close()},\texttt{read()},\texttt{write()},...
    \end{itemize}
    \item<3-> Can actually export \textit{any} file
    \item<4-> Crashes ungracefully when the client disconnects
  \end{itemize}
\end{frame}

\begin{frame}[c]{Protocol}
  \begin{center}
    \includegraphics[width=0.8\textwidth]{proto.png}
  \end{center}

  \begin{itemize}
    \item<1-> Structures passed through \texttt{INET} socket
    \item<2-> Raw data immediately follows when necessary
    \begin{itemize}
      \item<2-> \texttt{read()}, \texttt{write()}, ...
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\section{Protocol}

\begin{frame}[fragile]{Protocol}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{lstlisting}[gobble=8]
        enum FOPS {
          FOP_OPEN, FOP_RELEASE,
          FOP_READ, FOP_WRITE,
          _FOP_ERROR
        };

        struct fop_request {

          enum FOPS call;

          union {
            struct {
              int     flags;
              mode_t  mode;
            };

            size_t  count;
          };
        };
      \end{lstlisting}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{lstlisting}[gobble=8]
        struct fop_reply {

          enum FOPS call;

          union {
            int      open;
            int      close;
            ssize_t  read;
            ssize_t  write;
          };
        };
      \end{lstlisting}
    \end{column}
  \end{columns}
\end{frame}

\subsection{Client}

\begin{frame}[fragile]{Hello, World!}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \begin{lstlisting}[gobble=8]
        #include <linux/kernel.h>
        #include <linux/module.h>

        static int __init nc_init(void)
        {
          /* do stuff */
          return 0;
        }

        static int __exit nc_exit(void)
        {
          /* undo stuff */
        }

        MODULE_LICENSE("GPL");

        module_init(nc_init);
        module_exit(nc_exit);

      \end{lstlisting}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item initial setup
        \item cleanup important!
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]{Character Device Driver}
  \begin{lstlisting}[gobble=4]
    #include <linux/cdev.h>
    #include <linux/device.h>
    #include <linux/fs.h>

    static dev_t          nc_dev_t;
    static struct class*  nc_class;
    static struct cdev*   nc_cdev;
    static struct device* nc_device;

  \end{lstlisting}

  \begin{itemize}
    \item<1-> \texttt{dev\_t}: major/minor number
    \item<2-> \texttt{class}: subsystem
    \item<3-> \texttt{cdev}: handler for character devices
    \item<4-> \texttt{device}: userspace \texttt{/dev} node
    \item<5-> each has some form of registration mechanism
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Filesystem Handler}
  \begin{lstlisting}[gobble=4]
    #include <linux/fs.h>

    static int netchar_open(struct inode* inp, struct file* fp);
    static ssize_t netchar_read(struct file* fp, char* buffer,
                                size_t length, loff_t* offset);

    static struct file_operations nc_fops = {
      /* ... */
      .open  = nc_open,
      .read  = nc_read,
    }
  \end{lstlisting}

  \begin{itemize}
    \item<1-> required when registering character device
    \item<2-> buffer is a userspace pointer!
    \item<3-> event driven, blocking, asynchronous
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Socket Client}
  \begin{itemize}
    \item<1-> socket type: \texttt{AF\_INET}, \texttt{SOCK\_STREAM}, \texttt{IPPROTO\_TCP}
    \begin{itemize}
      \item<2-> should be \texttt{AF\_NETLINK}
      \item<2-> do not want to lose messages
    \end{itemize}
    \item<3-> no familiar \texttt{read()}, \texttt{write()}
  \end{itemize}
\end{frame}

\subsection{Server}

\subsection{Misc}

@@

% vim: set sw=2 ts=2 et spell : %
